Cluster Difference Imaging Photometric Survey (CDIPS) II. TOI$\,$837: A Validated Saturn-Sized Planet in IC~2602

### 2020-07-10 paper/ms.tex
update manuscript
Rhat = 1.0!

\begin{abstract}
  We report the discovery of TOI 837.01 and its validation as a
  transiting planet.  We characterize the system using data from the
  NASA TESS mission, the ESA Gaia mission, ground-based photometry,
  and spectroscopy from CHIRON, FEROS, and Veloce.  We find that
  TOI$\,$837 is a $T=9.9$ G0 dwarf in the open cluster IC 2602.  The
  star and validated planet are therefore $35^{+11}_{-5}$ million
  years old.  The planet is warm ($P = 8.3\,{\rm d}$) and slightly
  smaller than Jupiter.  Combining the photometry with a prior on the
  stellar parameters derived from the cluster color-magnitude diagram,
  we find that the transits are grazing with $b=0.951\pm0.012$ and
  $R_{\rm p} = 0.77 \pm 0.07 \,R_{\rm Jup}$.  From radial velocity
  monitoring, we limit the minimum mass to less than 1.20 $M_{\rm
  Jup}$ (3$\sigma$).  Grazing transits are a cause for concern, as
  they are usually indicative of astrophysical false positive
  scenarios.  Our follow-up data show that such scenarios are highly
  unlikely.  Our combined multi-color photometry, high-resolution
  imaging, and radial velocities rule out hierarchical eclipsing
  binary scenarios.  Background eclipsing binary scenarios, though
  limited by speckle imaging, formally remain a 0.2\% possibility.
  TOI 837.01 is therefore a validated adolescent exoplanet.  Further
  observations of its stellar obliquity, atmosphere, and mass should
  confirm its planetary nature beyond doubt, and could improve our
  understanding of how the physical and orbital properties of
  exoplanets change in time.
\end{abstract}

@@

Over the first 100 million years of their lives, exoplanet systems are
expected to undergo major physical and dynamical changes.  For a
typical Sun-like star, the protoplanetary disk disperses within
roughly 1--10 million years
\citep{mamajek_initial_2009,dullemond_inner_2010,williams_protoplanetary_2011}.
Gas giants presumably finish accreting before the end of disk
dispersal \citep{pollack_formation_1996}.  While rocky planets may
fully form within only a few million years
\citep{dauphas_hf-w-th_2011}, they can also undergo significant
growth over the next $\sim$10-100 million years through giant impacts
\citep[{\it
e.g.},][]{kleine_hf-w_2009,konig_earths_2011,morbidelli_building_2012,raymond_terrestrial_2014}.
The Moon, for instance, may have formed from debris ejected during a
collision between the proto-Earth and a planetesimal during Earth's
first 100 million years \citep{cameron_origin_1976,canup_origin_2001}.

@@

\begin{figure*}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=.85\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf TESS light curve of \tn\ from Sectors 10 and 11.} {\it Top}:
    \texttt{PDCSAP} median-subtracted relative flux at 2-minute
    sampling in units of parts-per-thousand ($\times 10^{-3}$).
    Starspot-induced variability is the dominant signal; flares
    are shown with gray crosses.  Dashed lines indicate the five
    transits observed by TESS.  {\it Bottom}: Zoomed windows of
    individual transits.  Gray lines are the best-fit transit model,
    which includes a local quadratic trend in each window.
		\label{fig:rawzoom}
	}
\end{figure*}

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=.47\textwidth]{f2a.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Phase-folded TESS transit.} Gray points are 2-minute PDCSAP flux
    measurements with the local spot-induced variation removed.  Black
    points are binned to 6-minute intervals. The black error-bar shows
    the median uncertainty for the black points.  The gray line is the
    best-fit model to the combined TESS and ground-based data
    (Table~\ref{tab:posterior}). 
		\label{fig:phasefold}
	}
\end{figure}

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=.47\textwidth]{f2b.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Ground-based follow-up photometry.} The data were acquired
    using the El Sauce 0.36m and ASTEP400 (0.40m). Black points are
    binned to 10-minute intervals.  The gray line is the best-fit
    model to the combined TESS and ground-based data. Models show
    $2\sigma$ lower limits on the transit depths in the Cousins-R and
    Johnson-B bandpasses used to rule out specific false positive
    scenarios (see Section~\ref{subsec:colorphot}).
		\label{fig:groundphot}
	}
\end{figure}

@@

Figure~\ref{fig:rawzoom} shows the data.  The dominant modulation
induced by starspots coming into and out of view has a peak-to-peak
amplitude of about 2.3\%, and a period of about 3 days.  The dips
suggestive of a grazing transiting planet recur roughly every 8 days,
and have a depth of about 0.4\%.  A few flares are visible.  A
phase-folded view of the TESS transits with starspot variability
removed is shown in Figure~\ref{fig:phasefold}. Our fitting procedure
for the models shown in these plots is described in
Section~\ref{subsec:planet}.

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.43\textwidth]{f5.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Scene of \tn.}
    {\it Top:} Mean TESS image of \tn\ over Sector~10, with a
    logarithmic grayscale. The yellow star is the position of \tn.
    White circles are neighboring stars with $T<16$ (brighter stars
    are larger). The black \texttt{X} and white \texttt{/} hatches
    show the apertures used to measure the background and target star
    flux, respectively. Dashed lines of constant declination and right
    ascension are shown.  {\it Bottom:} Digitized Sky Survey $R$-band
    image of the same field, with a linear grayscale.  Two stars of
    interest are ``Star A'' and ``Star B'', which were eventually
    excluded as being possible sources of the transits.
		\label{fig:scene}
	}
\end{figure}

@@

The Gaia imaging, reduced to its point-source catalog, provides
the initial context for analyzing the TESS data.  Stars brighter than
$T=16$, as queried from the Gaia DR2 source catalog, are shown with
white circles in Figure~\ref{fig:scene}, overlaid on the DSS2 plates
and TESS image.  Given its galactic latitude of $b=-6^\circ$, it is
not surprising that the field of \tn\ is crowded.  The resolved stars
that were of immediate concern for our false positive analysis were as
follows.
\begin{itemize}
  %
  \item \tn\ $\equiv$ TIC 460205581 ($T=9.9$). The target star.
  %
  \item Star A $\equiv$ TIC 847769574 ($T=14.6$), $2.3''$ west. The
    proper motions and parallax of this star imply it is comoving with
    \tn and is separated from it by $6.6\pm 0.1\,{\rm pc}$.
    It is therefore likely an \cn\ member, but unlikely to be a bound
    binary compnaion.
  %
  \item Star B $\equiv$ TIC 460205587 ($T=13.1$), $5.4''$ north.  The
    Gaia parallax implies this is a giant background star.
\end{itemize}
An additional source, TIC 847769581, is $4.9''$ from the target, but
too faint ($T=18.8$) to be the source of the observed transit signal.

The Gaia DR2 data for Star A seems poorly behaved.  While Star A has
$G=15.1$, and $Bp=14.9$, no $Rp$ magnitude is reported.
Correspondingly, no RUWE\footnote{ See the Gaia DPAC technical note
GAIA-C3-TN-LU-LL-124-01,
\url{http://www.rssd.esa.int/doc_fetch.php?id=3757412}, accessed
2020-06-22. } value is available.  We suspect that the photometric
failure to produce an $Rp$ magnitude, as well as its poor astrometric
fit, are likely due to blending with \tn.

@@

@@

We fitted the available time-series photometry of \pn\ as follows.
For the PDCSAP light curve, we first eliminated points that had
quality flags corresponding to any of bits $\{3, 4, 6, 8, 11, 12\}$.
This excluded cadences affected by coarse spacecraft pointing,
reaction wheel desaturation events, manual flags, cosmic ray hits, and
straylight from the Earth or Moon being present.  Inspecting the data,
we also manually excluded the two flares shown in
Figure~\ref{fig:rawzoom}.  We then trimmed the TESS data to windows of
$\pm 7\,{\rm hr}$ centered on each transit.

>>>>

\begin{abstract}
  We report the discovery of TOI 837.01 and its validation as a
  transiting planet.  We characterize the system using data from the
  NASA TESS mission, the ESA Gaia mission, photometry from El Sauce
  and ASTEP400,
  and spectroscopy from CHIRON, FEROS, and Veloce.  We find that
  TOI$\,$837 is a $T=9.9$ G0 dwarf in the open cluster IC 2602.  The
  star and validated planet are therefore $35^{+11}_{-5}$ million
  years old.  The planet is warm ($P = 8.3\,{\rm d}$) and slightly
  smaller than Jupiter.  Combining the transit photometry with a prior on the
  stellar parameters derived from the cluster color-magnitude diagram,
  we find that the transits are grazing with $b=0.93\pm0.01$ and
  $R_{\rm p} = 0.79 \pm 0.07 \,R_{\rm Jup}$.  From radial velocity
  monitoring, we limit the minimum mass to less than 1.20 $M_{\rm
  Jup}$ (3$\sigma$).  Grazing transits are a cause for concern, as
  they are usually indicative of astrophysical false positive
  scenarios.  Our follow-up data show that such scenarios are highly
  unlikely.  Our combined multi-color photometry, high-resolution
  imaging, and radial velocities rule out hierarchical eclipsing
  binary scenarios.  Background eclipsing binary scenarios, though
  limited by speckle imaging, formally remain a 0.2\% possibility.
  TOI 837.01 is therefore a validated adolescent exoplanet.  Further
  observations of its stellar obliquity, atmosphere, and mass should
  confirm its planetary nature beyond doubt, and could improve our
  understanding of how the physical and orbital properties of
  exoplanets change in time.
\end{abstract}

@@

Over the first 100 million years of their lives, exoplanet systems are
expected to undergo major physical and dynamical changes.  For a
typical Sun-like star, the protoplanetary disk disperses within
roughly 1--10 million years
\citep{mamajek_initial_2009,dullemond_inner_2010,williams_protoplanetary_2011}.
Gas giants presumably finish accreting before the end of disk
dispersal \citep{pollack_formation_1996}.  While rocky planets may
fully form within only a few million years
\citep{dauphas_hf-w-th_2011}, they can also undergo significant
growth over the next 10--100 million years through giant impacts
\citep[{\it
e.g.},][]{kleine_hf-w_2009,konig_earths_2011,morbidelli_building_2012,raymond_terrestrial_2014}.
The Moon, for instance, may have formed from debris ejected during a
collision between the proto-Earth and a planetesimal during Earth's
first 100 million years \citep{cameron_origin_1976,canup_origin_2001}.

@@

\begin{figure*}[!t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.79\textwidth]{f1.pdf}
		}

		\vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.77\textwidth]{f2a.pdf}
		}
	\end{center}
	\vspace{-0.6cm}
	\caption{
    {\bf TESS light curve of \tn.}
    {\it Top}:
    \texttt{PDCSAP} median-subtracted relative flux at 2-minute
    sampling in units of parts-per-thousand ($\times 10^{-3}$).
    Starspot-induced variability is the dominant signal; flares
    are shown with gray crosses.  Dashed lines indicate the five
    transits observed by TESS over Sectors 10 and 11.  {\it Middle}: Zoomed windows of
    individual transits.  Gray lines are the best-fit transit model,
    which includes a local quadratic trend in each window.
    {\it Bottom}: Phase-folded TESS transit. Gray points are 2-minute PDCSAP flux
    measurements with the local spot-induced variation removed.  Black
    points are binned to 6-minute intervals. The black error-bar shows
    the median uncertainty for the black points.  The gray line is the
    best-fit model to the combined TESS and ground-based data
    (Table~\ref{tab:posterior}). 
		\label{fig:tessphot}
	}
\end{figure*}

% \begin{figure*}[!t]
% 	\begin{center}
% 		\leavevmode
% 		\includegraphics[width=.85\textwidth]{f1.pdf}
% 	\end{center}
% 	\vspace{-0.7cm}
% 	\caption{
%     {\bf TESS light curve of \tn\ from Sectors 10 and 11.} {\it Top}:
%     \texttt{PDCSAP} median-subtracted relative flux at 2-minute
%     sampling in units of parts-per-thousand ($\times 10^{-3}$).
%     Starspot-induced variability is the dominant signal; flares
%     are shown with gray crosses.  Dashed lines indicate the five
%     transits observed by TESS.  {\it Bottom}: Zoomed windows of
%     individual transits.  Gray lines are the best-fit transit model,
%     which includes a local quadratic trend in each window.
% 		\label{fig:rawzoom}
% 	}
% \end{figure*}

% \begin{figure}[!t]
% 	\begin{center}
% 		\leavevmode
% 		\includegraphics[width=.47\textwidth]{f2a.pdf}
% 	\end{center}
% 	\vspace{-0.7cm}
% 	\caption{
%     {\bf Phase-folded TESS transit.} Gray points are 2-minute PDCSAP flux
%     measurements with the local spot-induced variation removed.  Black
%     points are binned to 6-minute intervals. The black error-bar shows
%     the median uncertainty for the black points.  The gray line is the
%     best-fit model to the combined TESS and ground-based data
%     (Table~\ref{tab:posterior}). 
% 		\label{fig:phasefold}
% 	}
% \end{figure}

@@

Figure~\ref{fig:tessphot} shows the data.  The dominant modulation
induced by starspots coming into and out of view has a peak-to-peak
amplitude of about 2.3\%, and a period of about 3 days.  The dips
suggestive of a grazing transiting planet recur roughly every 8 days,
and have a depth of about 0.4\%.  A few flares are visible.  A
phase-folded view of the TESS transits with starspot variability
removed is shown in Figure~\ref{fig:tessphot}. Our fitting procedure
for the models shown in these plots is described in
Section~\ref{subsec:planet}.

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f5.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Scene of \tn.}
    {\it Top:} Mean TESS image of \tn\ over Sector~10, with a
    logarithmic grayscale. The yellow star is the position of \tn.
    White circles are resolved Gaia sources with $T<16$ (brighter stars
    are larger). The black \texttt{X} and white \texttt{/} hatches
    show the apertures used to measure the background and target star
    flux, respectively. Dashed lines of constant declination and right
    ascension are shown.
    Two stars of interest are ``Star A'' and ``Star B'', which were eventually excluded as being possible sources of the transits.
    %{\it Bottom:} Digitized Sky Survey $R$-band
   % image of the same field, with a linear grayscale. 
		\label{fig:scene}
	}
\end{figure}

@@

The Gaia imaging, reduced to its point-source catalog, provides
the initial context for analyzing the TESS data.  Stars brighter than
$T=16$, as queried from the Gaia DR2 source catalog, are shown with
white circles in Figure~\ref{fig:scene}, overlaid on 
the TESS image.  Given its galactic latitude of $b=-6^\circ$, it is
not surprising that the field of \tn\ is crowded.  The resolved stars
that were of immediate concern for our false positive analysis were as
follows.
\begin{itemize}
  %
  \item \tn\ $\equiv$ TIC 460205581 ($T=9.9$). The target star.
  %
  \item Star A $\equiv$ TIC 847769574 ($T=14.6$), $2.3''$ west. The
    proper motions and parallax of this star imply it is comoving with
    \tn and is separated from it by $6.6\pm 0.1\,{\rm pc}$.
    It is therefore likely an \cn\ member, but unlikely to be a bound
    binary compnaion.
  %
  \item Star B $\equiv$ TIC 460205587 ($T=13.1$), $5.4''$ north.  The
    Gaia parallax implies this is a giant background star.
\end{itemize}
An additional source, TIC 847769581, is $4.9''$ from the target, but
too faint ($T=18.8$) to be the source of the observed transit signal.

The Gaia DR2 data for Star A seems poorly behaved.  While Star A has
$G=15.1$, and $Bp=14.9$, no $Rp$ magnitude is reported.
Correspondingly, no RUWE\footnote{ See the Gaia DPAC technical note
GAIA-C3-TN-LU-LL-124-01,
\url{http://www.rssd.esa.int/doc_fetch.php?id=3757412},
\texttt{2020-07-08}. } value is available.  We suspect that the photometric
failure to produce an $Rp$ magnitude, as well as its poor astrometric
fit, are likely due to blending with \tn.

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=.47\textwidth]{f2b.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Ground-based follow-up photometry.} The data were acquired
		using the 0.36m at El Sauce and the 0.40m ASTEP400 at Dome C. Black points are
		binned to 10-minute intervals.  The gray line is the best-fit
		model to the combined TESS and ground-based data. Models show
		$2\sigma$ lower limits on the transit depths in the Cousins-R and
		Johnson-B bandpasses used to rule out specific false positive
		scenarios (see Section~\ref{subsec:colorphot}).
		\label{fig:groundphot}
	}
\end{figure}

@@

We fitted the available time-series photometry of \pn\ as follows.
For the PDCSAP light curve, we first eliminated points that had
quality flags corresponding to any of bits $\{3, 4, 6, 8, 11, 12\}$.
This excluded cadences affected by coarse spacecraft pointing,
reaction wheel desaturation events, manual flags, cosmic ray hits, and
straylight from the Earth or Moon being present.  Inspecting the data,
we also manually excluded the two flares shown in
Figure~\ref{fig:tessphot}.  We then trimmed the TESS data to windows of
$\pm 7\,{\rm hr}$ centered on each transit.
